The landscape of chloroplast genome assembly tools

### 2019-06-05 bmc_article.tex
Update on Overleaf.

\author[
    addressref={aff2,aff3},
    noteref={n1},
    email={jan.freudenthal@uni-wuerzburg.de}
    ]{\inits{JAF}\fnm{Jan A} \snm{Freudenthal}}
\author[
    addressref={aff2},
    noteref={n1},
    email={simon-pfaff@gmx.de}
    ]{\inits{SP}\fnm{Simon} \snm{Pfaff}}
\author[
   addressref={aff1,aff3},
   email={niklas.terhoeven@uni-wuerzubrg.de}
]{\inits{NT}\fnm{Niklas} \snm{Terhoeven}}
\author[
   addressref={aff1},
   email={arthur.korte@uni-wuerzubrg.de}
]{\inits{AK}\fnm{Arthur} \snm{Korte}}
\author[
   addressref={aff2,aff1,aff3 },
   noteref={n2},
   email={markus.ankenbrand@uni-wuerzubrg.de}
]{\inits{MJA}\fnm{Markus J} \snm{Ankenbrand}}
\author[
   addressref={aff1,aff4},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   noteref={n2},                        % id's of article notes, if any
   email={frank.foerster@uni-wuerzburg.de}   % email address
]{\inits{FF}\fnm{Frank} \snm{Förster}}

@@

\address[id=aff2]{%                           % unique id
  \orgname{Department of Bioinformatics, University of W\"{u}rzburg}, % university, etc
  \street{Biozentrum, Am Hubland},                     %
  \postcode{97074},                                % post or zip code
  \city{W\"{u}rzburg},                              % city
  \cny{Germany}                                    % country
}
\address[id=aff1]{%
  \orgname{Center for Computational and Theoretical Biology, University of W\"{u}rzburg},
  \street{Campus Hubland Nord},
  \postcode{97074}
  \city{W\"{u}rzburg},
  \cny{Germany}
}
\address[id=aff3]{%
  \orgname{AnaLife Data Science},
  \street{Friedrich-Bergius-Ring 15},
  \postcode{97076}
  \city{W\"{u}rzburg},
  \cny{Germany}
}
\address[id=aff4]{%
  \orgname{Fraunhofer IME-BR},
  \street{Winchester Str.\ 2},
  \postcode{97076}
  \city{Gie\ss{}en},
  \cny{Germany}
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract
The special issue will accept manuscripts presenting:

- in-depth analyses of advantages

-limitations of state-of-the-art methods

-demonstration on a variety of representative datasets and simulations

-clear guidelines/recommendations for the end-user

-insight into future directions for methods
\parttitle{First part title} %if any
Text for this section.

\parttitle{Second part title} %if any
Text for this section.
\end{abstract}

@@

%%%%%%%%%%%%%%%%
%% Background %%
%%
\newpage
\section*{Introduction}
\subsection*{General introduction and motivation}
The chloroplast as the photosynthetic organelle harboring its own genome has been an early target for sequencing.\ak{maybe even 1-2 sentences about general chloroplast importance}
The first chloroplast genome sequences were obtained as early as 1986 \cite{ohyama_chloroplast_1986,shinozaki_complete_1986}.
These efforts elucidated the general genome organization and structure of chloroplasts, expanded by many more studies over the years.
Chloroplast genome content and structure are reviewed elsewhere \cite{wicke_evolution_2011,green_chloroplast_2011}.
Chloroplast genomes are widely used for evolutionary analyses \cite{martin_plastid_2010,xiao-ming_inferring_2017}, barcoding \cite{kress_use_2005,hollingsworth_dna_2009,de_vere_dna_2015}, and meta-barcoding \cite{bell_review_2016,deiner_environmental_2017}.
Interesting aspect of chloroplast genomes are their reduced size through endosymbiotic gene transfer \cite{martin_evolutionary_2002,timmis_endosymbiotic_2004} and the strong conservation of the remaining genes (cite).
Despite the overall high conservation there are striking differences between groups (e.g. loss of whole gene families like \genename{ndh} genes in Droseraceae \cite{nevill_plastome-wide_2019} and extreme evolutionary cases with very low GC content and a modified genetic code \cite{su_novel_2019}).
These differences call for comparative genomic approaches.

It is much easier to decipher the complete chloroplast genome than the complete core genome, due its comparably smaller size. For example \textit{Arabidopsis thaliana} core genome is approximately \SI{125}{\mega\basepair} in length \cite{schmuths2004,cao2011} while the size of the chloroplast genome with \SI{154}{\kilo\basepair} is more than \SI{800}{\times} smaller than the core genome \cite{sato1999}.    
However, the specific genome structure (esp. the large inverted repeat) complicates automated resolution with short read technologies (cite?).
Many genome sequencing projects contain chloroplast reads as by-product.
In some cases the chloroplast data is even considered contamination so experimental protocols for reducing their content have been developed \cite{lutz_isolation_2011}.
The resolved chloroplast genome helps to remove these reads, which can improve assembly of the core genome.
Databases exist, containing short read data for species where no reference chloroplast is publicly available.
There is a lot of potential use cases for chloroplast genomes \cite{tonti-filippini_what_2017}
Resolving these (hidden treasures) would allow for more and larger scale comparative studies on whole chloroplasts.
In a different application researchers use shallow genome sequencing to reconstruct full chloroplast genomes for use as super-barcodes \cite{coissac_barcodes_2016}.
There are also applications in biotechnology and genetic engineering \cite{daniell_chloroplast_2016}.

\subsection*{Chloroplast genome architecture}
Maybe separate section describing a prototypical chloroplast: LSC, IRa, SSC, IRb. Gene number and content. Size ranges. Evolutionary history incl. endosymbiontic gene transfer. Reduced genomes in parasitic plants. Loss of \genename{ndh} gene family (maybe too specific). Preliminary SNP analysis in \plant{At} to estimate intraspecific variability? Is there anything known about intra-organismic variability (do some plants have multiple populations of chloroplasts? \ak{Yes, they do , Heteroplasmy} probably off topic).\ak{nope, I think this is important}

@@

In general, obtaining a chloroplast genome from WGS data requires two steps. First, the chloroplast reads have to be extracted from the mixed sequencing data. 
The second step is the assembly and resolving the special circular structure including the inverted repeat.
The extraction of the reads can be achieved by mapping the reads to a reference chloroplast. \todo{example}
This can also be done iteratively, as done by \todo{\ioga{}?}.
A different approach relies on the higher coverage of chloroplast data in the whole genome sequencing data set. A kmer analysis can be used here to extract the most frequent reads. An example for this is implemented in \ce{}.
A third method, solving both steps simultaneously, uses a reference chloroplast as seed and then assembles the reads based on that. \todo{example}

Once the chloroplast reads are extracted from the data set, they need to be assembled. To resolve the special structure, it is possible to analyze the assembly graph and infer the circular structure and the inverted repeat from that. This approach is used in \ce{}. A second method uses local alignments to the assembly to find the correct order of the contigs. \todo{example}

\todo{cite twyford strategies 2017}

%\todo{ausbauen}
%1. Getting chloroplast reads out of mixed files
%Mapping (potentially iterative), k-mer coverage analysis, seeded assembly
%2. Assembly of chloroplast (incl special structure with IRs)
%normal assembly - then blast, assembly then graph resolution, might be combined with step 1.

\subsection*{Purpose and scope of this study}
The goal of this study is to compare the effectiveness and efficiency of existing open source command-line tools to de-novo assemble whole chloroplast genomes from raw mixed genomic data sets with minimal configuration.
This means: no data preparation, no specific reference (but \plant{At} if required), default parameters, no manual finishing.
We further restricted our benchmark to paired end Illumina data sets as these are the most common sequencing platform \cite{Goodwin2016}. % done by Frank \todo{tools can not handle passt hier finde ich nicht als Begründung. Wir geben gerade vor was uns interessiert und wählen später Tools aus die dazu passen. Sonst gibt es einen Ringschluss zwischen Scope und Tool Selection.}

In our opinion this reflects the most common use cases: (1) a user trying a tool quickly without digging into options for fine tuning and (2) large scale automatic applications.
Still, we acknowledge that the performance of the tools might be significantly improved by optimizing parameters (and references if applicable) for each data set specifically.
(However, this was outside the scope of this study.)

%Clearly state criteria. Work with mixed data. Full automation. De novo - no manual selection of appropriate reference. No manual finishing. The step we are interested in is: from raw paired Illumina (!) reads to a fasta file containing the assembly (optimally full, circularized and on one contig). We are not interested in annotation or any additional contigs (genomic, mitochondrion, rRNA, etc.).
%\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}

\section*{Methods}
Source code for all methods is available at \url{https://github.com/chloroExtractorTeam/benchmark} and archived under \zenodorepo{}.
All docker images are published on \url{https://cloud.docker.com/u/chloroextractorteam/} and are named with a leading \texttt{benchmark\_} (\cref{tab:dockerimages}).
Our simulated data sets are stored at \zenododataset{}.
This study adheres to the guidelines for computational method benchmarking \cite{weber_essential_2018}.

\subsection*{Tool Selection}

We included tools designed for assembling chloroplasts from whole genome paired end Illumina sequencing data. The tools must be available as open source software and allow execution via a command line interface. A graphical user interface is not suitable for
running automated comparisons. The following tools were determined to be within the scope of this study:

\oa{} \cite{coissac_barcodes_2016}, 
\ce{} \cite{j_ankenbrand_chloroextractor:_2018}, 
\fp{} \cite{mckain__fast-plast_2017}, 
\ioga{} \cite{bakker_herbarium_2016}, 
\np{} \cite{dierckxsens_novoplasty:_2017}, 
\go{} \cite{jin_getorganelle:_2018}, 
\cassp{} \cite{sancho_comparative_2018}.

Other related tools for assembling chloroplasts but outside the scope of this study are \texttt{Organelle PBA} \cite{Soorni2017}, \texttt{sestaton/Chloro} \cite{sestaton}, \texttt{Norgal}  \cite{Al-Nakeeb2017} and \texttt{MitoBim} \cite{mitobim2013}.

\texttt{Organelle PBA} is designed for PacBio data and does not work with paired Illumina data only.
\texttt{sestaton/Chloro} fits all criteria but is flagged as a work in progress and development seems to have ended two years ago.
\texttt{Norgal} is a tool to extract organellar DNA from whole genome data based on a k-mer frequency approach. However the final output is a set of contigs of mixed mitochondrial and plastid origin. The suggested approach to get a finished chloroplast genome is to run \np{} on the ten longest contigs.
\texttt{MitoBim} is specifically designed for mitochondrial genomes. Even though there is a claim by the author that it can be used for chloroplasts as well (\url{https://github.com/chrishah/MITObim/issues/16}, accessed: 2019-05-27), there is no further description on how to do that. %\todo{haben wir es versucht? - nein}
Additionally, there is a protocol for the \texttt{Geneious} \cite{geneious} software available \cite{geneious-protocol}.
However, \texttt{Geneious} is closed source and GUI based, which is not in the scope of this study.
There is also another publication describing a method for assembling chloroplasts \cite{method-description-paper}. However, the link to the software is not active anymore.

\subsection*{Our Setup}
% done by Frank \todo{ausformulieren}
We want to use accomplish a minimum of different parameter settings for all assembly programs.
Therefore, we decided to specify that all programs have to work based on two input files, representing a data set's forward (\texttt{forward.fq}) and reverse input (\texttt{reverse.fq}).
Moreover, depending on the assembler, files with different names and locations are generated as final output.
Those different files are copied and renamed to ensure \texttt{output.fa} as exclusive output of an assembly approach.
Additionally, we want to an environment variable to control the number of allowed threads.
All three requirements (defined input file names, defined output file name, thread number control via environment variable) are ensured by a simple wrapper script (\texttt{wrapper.sh}).
Finally, for a maximum of reproducibility all programs are bundled into individual docker images based on a central base image which provides all the required software.
That docker images were used for the recording of the consumption of computational resources.
Furthermore, all our docker images are converted into singularity containers for the quantitative measurement on simulated and real data sets .
%
Singularity container were built from docker images for usage on Julia-HPC of JMU using Singularity v.2.5.2 \cite{kurtzer2017singularity}
All singularity containers were run on the Julia-HPC on Intel® Xeon® Gold \num{6140} Processors using slurm workload manager version 17.11.8 \cite{Jette02slurm}. Assemblies run on \num{4} threads using \SI{10}{\gibi\byte} RAM with a timelimit of \SI{48}{\hour}. 
\subsection*{Data}
\subsubsection*{Simulated}
To avoid suffering from sequencing errors and biological variances, we simulated perfect reads based on the \plant{At} (TAIR10) chloroplast assembly \cite{tair10}.
We used a sliding window approach with \texttt{seqkit} \cite{seqkit}. The exact commands are documented in \href{https://github.com/chloroExtractorTeam/benchmark/blob/master/03_representative_datasets.md}{\texttt{03\_representative\_datasets.md}} in \zenododataset{}.
For the final simulated data sets reads based on pseudo mixtures of TAIR10's core and chloroplast genome:  \num{0}:\num{1}, \num{1}:\num{10}, \num{1}:\num{100}, and \num{1}:\num{1000}.
Additionally, we generated different read lengths (\SIlist{150;250}{\basepair}) and subsets with exactly 2 million read pairs.

@@

\subsubsection*{Real}
Additionally, we selected real datasets deposited at SRA \cite{sra2010}.
Among all datasets that matched this search \texttt{((((((("green plants"[orgn]) AND "wgs"[Strategy]) AND "illumina"[Platform]) AND "biomol dna"[Properties]) AND "paired"[Layout]) AND "random"[Selection])) AND "public"[Access]} (on 2019-03-28) we selected one dataset per species with a reference chloroplast in CpBase \cite{cpbase}.
In total, this accumulated to \num{380} datasets, including \num{6} without data and \num{5} with malformatted data \todo{soll ich da die genauen accessions nochmal raussuchen? bzw. im supp table flaggen?}. This leads to a total set of \num{369} datasets \todo{see supplementary table - da die real\_datasets.tsv, die sraruninfo oder eine custom table?}.

\subsection*{Evaluation Criteria}
\subsubsection*{Computational Resources}
Due to computational resources are limited, we evaluated the effect of different input parameters.
To monitor the influence, we recorded the mean and the peak CPU usage, the peak memory consumption, and the size of the assembly folder. 
As input data, we used different data sets comprising \numlist{25000;250000;2500000} read pairs based on our simulated perfect reads.
Therefore, the simulated perfect read files have been read repeatedly to achieve the required number of read pairs.
We used our docker image setup (\cref{tab:dockerimages}) to run all assembly programs three times on a given parameter setting.
The setting combined the different input data and different number of threads to use (\numlist{1;2;4;8}).
Some programs used more CPU threads than specified, therefore, the number of CPUs available have been fixed using the CPUs option while running the docker run command.
For each assembly setting, we recorded the peak memory consumption, the CPU usage for calculation of mean and peak CPU usage, and the size of the folder where the assembly was calculated.
The values of CPU and memory usage have been obtained by docker.
The disk usage was estimated using the GNU tool \texttt{du}.
We used \texttt{GNU parallel} for queuing of the different settings \cite{Tange2011a}.

\subsubsection*{Qualitative}

The qualitative evaluation is mainly based on the reviewer guidelines for the Journal of Open Source Software (JOSS) \cite{joss}.
To create a standard environment, all tools were tested in a fresh default installation of Ubuntu 18.04.2 running in a virtual machine (VirtualBox Version 5.2.18\_Ubuntu r123745). We chose this setup instead of the docker container, because it resembles a typical user environment better than the minimal docker installation. The tools were installed according to their installation instructions and
the provided tutorial or example usage was executed. During the evaluation, the following
questions were asked:

\begin{itemize}
    \item Is the tool easy to install? 
    \item Is there a way to test the installation or a tutorial on how to use the tool? 
    \item Is there a good documentation on the parameter settings? 
    \item Is the tool maintained (issues answered, implementation of new features)? 
    \item Is the tool Open Source?
\end{itemize}

>>>>

\author[
    addressref={aff1,aff3},
    noteref={n1},
    email={jan.freudenthal@uni-wuerzburg.de}
    ]{\inits{JAF}\fnm{Jan A} \snm{Freudenthal}}
\author[
    addressref={aff1,aff2},
    noteref={n1},
    email={simon-pfaff@gmx.de}
    ]{\inits{SP}\fnm{Simon} \snm{Pfaff}}
\author[
   addressref={aff1,aff3},
   email={niklas.terhoeven@uni-wuerzubrg.de}
]{\inits{NT}\fnm{Niklas} \snm{Terhoeven}}
\author[
   addressref={aff1},
   email={arthur.korte@uni-wuerzubrg.de}
]{\inits{AK}\fnm{Arthur} \snm{Korte}}
\author[
   addressref={aff1,aff3,aff2 },
   noteref={n2},
   email={markus.ankenbrand@uni-wuerzubrg.de}
]{\inits{MJA}\fnm{Markus J} \snm{Ankenbrand}}
\author[
   addressref={aff2,aff4},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   noteref={n2},                        % id's of article notes, if any
   email={frank.foerster@uni-wuerzburg.de}   % email address
]{\inits{FF}\fnm{Frank} \snm{Förster}}

@@

\address[id=aff1]{%
  \orgname{Center for Computational and Theoretical Biology, University of W\"{u}rzburg},
  \street{Campus Hubland Nord},
  \postcode{97074}
  \city{W\"{u}rzburg},
  \cny{Germany}
}

\address[id=aff3]{%
  \orgname{AnaLife Data Science},
  \street{Friedrich-Bergius-Ring 15},
  \postcode{97076}
  \city{W\"{u}rzburg},
  \cny{Germany}
}
\address[id=aff2]{%                           % unique id
  \orgname{Department of Bioinformatics, University of W\"{u}rzburg}, % university, etc
  \street{Biozentrum, Am Hubland},                     %
  \postcode{97074},                                % post or zip code
  \city{W\"{u}rzburg},                              % city
  \cny{Germany}                                    % country
}

\address[id=aff4]{%
  \orgname{Fraunhofer IME-BR},
  \street{Winchester Str.\ 2},
  \postcode{97076}
  \city{Gie\ss{}en},
  \cny{Germany}
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstractbox}
\begin{abstract} % abstract
%One or two sentences providing a basic introduction to the field, comprehensible to a scientist in any discipline.
Chloroplasts are photosynthetic active organelles in plant cells and contain their own genomic information.
That genome can be utilized in different scientific fields like phylogenetics or biotechnology.
%Two to three sentences ofmore detailed background, comprehensible to scientists in related disciplines.
Thus, different assemblers have been developed specialized in chloroplast assemblies.
Those assemblers often use the output of whole genome sequencing experiments as input.
Such sequencing data contain usually the complete chloroplast genome information, even if the sequencing aims for the core genome.
%One sentence clearly stating the general problem being addressed by this particular study.
The Different assembly tools have never been systematically compared. %to each other concerning success rate or computational requirements.
%One sentence summarising the main result (with the words “here we show” or their equivalent).
Here we present a benchmark of seven chloroplast assembly tools, which are under active\todo{Niklas: da kann man bei IOGA drüber streiten} development and are capable to succeed in more than \SI{60}{\percent} of real data sets.
%Two or three sentences explaining what the main result reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.
Our results show significant differences between the tested assemblers in terms of generating whole chloroplast genome sequences and computational requirements.
Moreover, we suggest further development to improve user experience and success rate.
%One or two sentences to put the results into a more general context. 
In terms of reproducibility, we created docker images for each tested tool, which are available for the scientific community.
Following the presented guidelines, users are able to analyse and screen data sets for chloroplast genomes using only standard computer infrastructure.
%Two or three sentences to provide a broader perspective, readily comprehensible to a scientist in any discipline, may be included in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced by their inclusion. 
Thus large scale screening for chloroplasts as hidden treasures within genomic sequencing data is feasible.

%\ff{Abstract}
%Useful as always: \url{https://cbs.umn.edu/sites/cbs.umn.edu/files/public/downloads/Annotated_Nature_abstract.pdf}
%The special issue will accept manuscripts presenting:
%
%- in-depth analyses of advantages
%
%-limitations of state-of-the-art methods
%
%-demonstration on a variety of representative datasets and simulations
%
%-clear guidelines/recommendations for the end-user
%
%-insight into future directions for methods
%\parttitle{First part title} %if any
%Text for this section.
%
%\parttitle{Second part title} %if any
%Text for this section.
\end{abstract}

@@

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Introduction}
\subsection*{General introduction and motivation}
Chloroplasts are essential organelles present in plant cells and the cells of some protists. Chloroplasts enable the conversion of light energy into chemical energy via photosyntesis. They harbor their own ribosomes and a circular DNA genome with a size between \SIrange{120}{160}{\kilo\basepair} \cite{palmer_1985}.
Because of this small size, the chloroplast genome has been an early target for sequencing.
The first chloroplast genome sequences were obtained as early as 1986 \cite{ohyama_chloroplast_1986,shinozaki_complete_1986}.
These early efforts elucidated the general genome organization and structure of the chloroplast DNA. 
Chloroplast genome content and structure are reviewed for example in \cite{wicke_evolution_2011,green_chloroplast_2011}.
Chloroplast genomes are widely used for evolutionary analyses \cite{martin_plastid_2010,xiao-ming_inferring_2017}, barcoding \cite{kress_use_2005,hollingsworth_dna_2009,de_vere_dna_2015}, and meta-barcoding \cite{bell_review_2016,deiner_environmental_2017}.
Interesting aspects of chloroplast genomes are their small size (\SIrange{120}{160}{\kilo\basepair},\cite{palmer_1985}), caused through endosymbiotic gene transfer \cite{martin_evolutionary_2002,timmis_endosymbiotic_2004} and the low number of \numrange{100}{120} genes that are still encoded on the chloroplast genome \cite{wicke_evolution_2011}.
Despite the overall high conservation of the genome sequence, there are striking differences in the gene content between different groups (e.g. the loss of the whole \genename{ndh} gene family in Droseraceae \cite{nevill_plastome-wide_2019}). Even more extreme evolutionary cases, where chloroplasts show a very low GC content and a modified genetic code are described \cite{su_novel_2019}.

These differences call for comparative genomic approaches.
Given the small size, it is much easier to decipher the complete chloroplast genome than the complete core genome. For example the \plant{At} core genome is approximately \SI{125}{\mega\basepair} in length \cite{schmuths2004,cao2011} while the size of the \plant{At} chloroplast genome with \SI{154}{\kilo\basepair} is more than \SI{800}{\times} smaller \cite{sato1999}.

Even if only a single chloroplast is located inside a plant cell, several hundreds copies of the chloroplast genome exists in each cell \cite{kumar_2014,bendich_1987}.
Therefore, many genome sequencing projects contain chloroplast reads as by-product.
In some cases the chloroplast data is even considered contamination and experimental protocols for reducing their content have been developed \cite{lutz_isolation_2011}.
An alternative approach to improve the assembly of the core genome would be to first resolve the chloroplast genome and afterwards use this information to remove those reads that map to the chloroplast genome.

Structurally, two inverted repeats (IR\textsubscript{A} and IR\textsubscript{B}) of \SIrange{10}{76}{\kilo\basepair} divide the chloroplast genome into a large (LSC) and a small single copy (SSC) region \cite{palmer_1985}.
Those large inverted repeats complicates automated resolution with short read technologies\cite{Wang2018}.
Moreover, the existence of different chloroplasts within a single individual, and thus multiple different chloroplast genomes, have been described for different plants \cite{corriveau_1988,Chat2002,Scar2016}.
Although the origin and evolutionary importance of this phenomena, - called heteroplasmy - are only poorly understood, the assembly of whole chloroplast genomes might be hindered.

Databases containing short read data for species where no reference chloroplast sequence is publicly available, exist.
The availability of whole chloroplast genomes would enable large scale comparative studies \cite{tonti-filippini_what_2017}.
Additionally, reconstructed full chloroplast genomes have been used as super-barcodes \cite{coissac_barcodes_2016}, for biotechnology applications and genetic engineering \cite{daniell_chloroplast_2016}.

%\subsection*{Chloroplast genome architecture}\ff{Chloroplast genome architecture section}
%
%Maybe separate section describing a prototypical chloroplast: LSC, IRa, SSC, IRb. Gene number and content. 
%Size ranges. 
%Evolutionary history incl. endosymbiontic gene transfer. 
%Reduced genomes in parasitic plants. Loss of \genename{ndh} gene family (maybe too specific).
%Preliminary SNP analysis in \plant{At} to estimate intraspecific variability? Is there anything known about intra-organismic variability (do some plants have multiple populations of chloroplasts? \ak{Yes, they do , Heteroplasmy} probably off topic).\ak{nope, I think this is important}

@@

Different strategies have been developed to assemble chloroplast genomes \cite{twyford_strategies_2017}.
In general, obtaining a chloroplast genome from WGS data requires two steps. First, the chloroplast reads have to be extracted from the mixed sequencing data. 
The second step is the assembly and resolving the special circular structure including the inverted repeats.
The extraction of the reads can be achieved by mapping the reads to a reference chloroplast. \cite{Vinga2012}.
A different approach that does not perform alignments, relies on the higher coverage of chloroplast data in the whole genome sequencing data set\cite{Chan2013}. Here, a $k$-mer analysis can be used to extract the most frequent reads. An example for this is implemented in \ce{}.
A third method combines both approaches by useing a reference chloroplast as seed and simultaneously assembling the reads based on $k$-mers \cite{dierckxsens_novoplasty:_2017}.

%\todo{Niklas: wir nennen hier immer nur \ce{} als beispiel - wir sollten auch andere nennen}
%\todo{example}
%\ak{e.g. HomBlocks: A multiple-alignment construction pipeline for organelle phylogenomics based on locally collinear block searching, aber ist nicht inkluded, also können wir zitieren, müssen aber nicht}

\subsection*{Purpose and scope of this study}
The goal of this study is to compare the effectiveness and efficiency of existing open source command-line tools to de-novo assemble whole chloroplast genomes from raw genomic data sets with minimal configuration. 
This includes no need for extensive data preparation, no need for a specific reference (apart from \plant{At}), no need to change default parameters, no manual finishing.
We further restricted our benchmark to paired end Illumina data sets as these are routinely generated by modern sequencing platforms \cite{Goodwin2016}.

In our opinion this reflects the most common use cases: (1) a user trying a tool quickly without digging into options for fine tuning and (2) large scale automatic applications.
Still, we acknowledge that the performance of the tools might be significantly improved by optimizing parameters (and references if applicable) for each data set specifically. However an exhaustive comparison - including tuning of all different possible parameters for each tool- would not been possible within the scope of this study.

Our results will enable the discovery of novel chloroplast genomes as well as an assembly of inter/intra-individual differences in the respective chloroplast genomes.

\section*{Methods}
\subsection*{Data availability}
Source code for all methods used is available at \url{https://github.com/chloroExtractorTeam/benchmark} and archived in zenodo under \zenodorepo{}.
All docker images are published on \url{https://cloud.docker.com/u/chloroextractorteam/} and are named with a leading \texttt{benchmark\_} (\cref{tab:dockerimages}).

To enable a fair comparison of all tools, we generated simulated sequencing data.
Those simulated data sets are stored at \zenododataset{}.
This study adheres to the guidelines for computational method benchmarking \cite{weber_essential_2018}.

\subsection*{Tool Selection}
We included tools designed for assembling chloroplasts from whole genome paired end Illumina sequencing data. As a requirement, all tools must be available as open source software and allow execution via a command line interface. As a graphical user interface is not suitable for
running automated comparisons, tools only providing a graphical interface have not been included. The following tools were determined to be within the scope of this study:
\oa{} \cite{coissac_barcodes_2016}, 
\ce{} \cite{j_ankenbrand_chloroextractor:_2018}, 
\fp{} \cite{mckain__fast-plast_2017}, 
\ioga{} \cite{bakker_herbarium_2016}, 
\np{} \cite{dierckxsens_novoplasty:_2017}, 
\go{} \cite{jin_getorganelle:_2018}, and
\cassp{} \cite{sancho_comparative_2018}.

Other related tools for assembling chloroplasts that did not meet our criteria and are therefore outside the scope of this study are for example:
\texttt{Organelle PBA} \cite{Soorni2017}, \texttt{sestaton/Chloro} \cite{sestaton}, \texttt{Norgal}  \cite{Al-Nakeeb2017}, and \texttt{MitoBim} \cite{mitobim2013}.

\texttt{Organelle PBA} is designed for PacBio data and does not work with paired Illumina data alone.
\texttt{sestaton/Chloro} fits our criteria, but it is flagged as work in progress and development and support seem to have ended two years ago.
\texttt{Norgal} is a tool to extract organellar DNA from whole genome data based on a \textit{k}-mer frequency approach. However the final output is a set of contigs of mixed mitochondrial and plastid origin. The suggested approach to get a finished chloroplast genome is to run \np{} on the ten longest contigs. Therefore we only included \np{} with the default settings and excluded \texttt{Norgal}.
\texttt{MitoBim} is specifically designed for mitochondrial genomes. Even though there is a claim by the author that it can be used for chloroplasts as well, there is no further description on how to do that \cite{mitobim_issue16}.

Additionally, there is a protocol for the \texttt{Geneious} \cite{geneious} software available \cite{geneious-protocol}.
However, \texttt{Geneious} is closed source and GUI based, which is not in the scope of this study.
There is also another publication describing a method for assembling chloroplasts \cite{method-description-paper}. However, the link to the software is not active anymore.

\subsection*{Our Setup}
% done by Frank \todo{ausformulieren}
We want to use accomplish a minimum of different parameter settings for all assembly programs to enable a fair comparison. 
Therefore, we decided to specify that all programs have to work based on two input files, representing a data set's forward (\texttt{forward.fq}) and reverse  (\texttt{reverse.fq}) sequence file in FASTQ format.
Depending on the assembler, output files with different names and locations are generated.
Those different files are copied and renamed to ensure that each assembly approach produces the same output file (\texttt{output.fa}). 
Additionally, we set an environment variable for all programs to control the number of allowed threads.
All three requirements (defined input file names, defined output file name, thread number control via environment variable) are ensured by a simple wrapper script (\texttt{wrapper.sh}).
Finally, for a maximum of reproducibility all programs have been bundled into individual docker images based on a central base image which provides all the required software. Those docker images were used for the recording of the consumption of computational resources.
Furthermore, all our docker images have been converted into singularity containers for the quantitative measurement on simulated and real data sets.
Singularity container were built from docker images for usage on a HPC-environment using Singularity v.2.5.2 \cite{kurtzer2017singularity}
All singularity containers were run on %the Julia-HPC 
\ak{I would note refer to Julia in the ms, maybe we could add it as a link}
Intel® Xeon® Gold \num{6140}  Processors using a Slurm workload manager version 17.11.8 \cite{Jette02slurm}. Assemblies were run on \num{4} threads using \SI{10}{\gibi\byte} RAM with a time limit of \SI{48}{\hour}. 
\subsection*{Data}
\subsubsection*{Simulated}
To avoid suffering from sequencing errors and biological variances, we simulated perfect reads based on the \plant{At} (TAIR10) chloroplast assembly \cite{tair10}.
We used a sliding window approach with \texttt{seqkit} \cite{seqkit}. The exact commands are documented in \href{https://github.com/chloroExtractorTeam/benchmark/blob/master/03_representative_datasets.md}{\texttt{03\_representative\_datasets.md}} in \zenododataset{}.
For the final simulated data sets reads based on mixtures of the \plant{At} (TAIR10) core and chloroplast genome were generated with different ratios ( \num{0}:\num{1}, \num{1}:\num{10}, \num{1}:\num{100}, and \num{1}:\num{1000}).
Additionally, we generated data with different read lengths (\SIlist{150;250}{\basepair}). All data simulated contain exactly 2 million read pairs.

@@

\subsubsection*{Real}
We selected real data deposited at SRA \cite{sra2010}.
We searched all data that matched  \texttt{((((((("green plants"[orgn]) AND "wgs"[Strategy]) AND "illumina"[Platform]) AND "biomol dna"[Properties]) AND "paired"[Layout]) AND "random"[Selection])) AND "public"[Access]} \cite{sra_search_term}. 
For each species with a reference chloroplast in CpBase \cite{cpbase}, we selected one data set of those.
In total, this accumulated to \num{369} data sets (see supplement for details).

\subsection*{Evaluation Criteria}
\subsubsection*{Computational Resources}
%Due to computational resources are limited, we evaluated the effect of different input parameters.
We recorded the mean and the peak CPU usage, the peak memory consumption, and the size of the assembly folder for each program. 
As input data, we used different data sets comprising \numlist{25000;250000;2500000} read pairs sampled from  our simulated reads.
%Therefore, the simulated perfect read files have been read repeatedly to achieve the required number of read pairs.
%\ak{etwas umständlich formuliert, hoffe ich hab es jetzt richtig formuliert und verstanden}
We used our docker image setup (\cref{tab:dockerimages}) \ak{check NUmmerierung tables, oben war auch schon table 2} to run all assembly programs three times for each parameter setting.
The different settings combined different input data and different number of threads to use (\numlist{1;2;4;8}).
%\ak{in 'Our Setup' before, schreiben wir 10GiB und 4 threads, müsste dann vielleicht geändert werden}
Some programs will use more CPU threads than specified, therefore, the number of CPUs available have been fixed using the CPU option while running the docker run command.
For each assembly setting, we recorded the peak memory consumption, the CPU usage (mean and peak CPU usage) and the size of the folder where the assembly was calculated.
The values of CPU and memory usage have been obtained by docker.
The disk usage was estimated using the GNU tool \texttt{du}.
We used \texttt{GNU parallel} for queuing of the different settings \cite{Tange2011a}.

\subsubsection*{Qualitative}
The qualitative evaluation is mainly based on the reviewer guidelines for the Journal of Open Source Software (JOSS) \cite{joss}.
To create a standard environment, all tools were tested in a fresh default installation of Ubuntu 18.04.2 running in a virtual machine (VirtualBox Version 5.2.18\_Ubuntu r123745). We chose this setup instead of the docker container, because it resembles a typical user environment better than the minimal docker installation. The tools were installed according to their installation instructions and
the provided tutorial or example usage was executed. During the evaluation, the following questions were asked:

\begin{itemize}
    \item Is the tool easy to install? 
    \item Is there a way to test the installation or a tutorial on how to use the tool? 
    \item Is there a good documentation on the parameter settings? 
    \item Is the tool maintained (issues answered, implementation of new features)? 
    \item Is the tool Open Source?
\end{itemize}
\newline
These questions were answered with \good{}, \ok{} or \bad{}, depending on the quality of the result. 
For example, a \good{} installation utilizes a automated package or dependency management like \texttt{apt},
CRAN, docker, etc. An \ok{} installation procedure provides a custom script to install everything or at least lists all dependencies. A \bad{} installation procedure fails to list important dependencies or produces errors, that prevent a successful installation without exhaustive debugging.

After an initial evaluation, we contacted all authors via their GitHub or GitLab issue tracking to communicate potential flaws we found.
